% Chapter Template

\chapter{Activity - Rotation Investigation} % Main chapter title

\label{Chapter5} 

\epigraph{\itshape Sometimes the wheel turns slowly, but it turns.}{---Lorne Michaels}

\section{Introduction}

The results presented in Chapters \ref{Chapter3} and \ref{Chapter4} investigated the age-activity relationship through coronal and chromospheric emission. However, the age-activity relationship is a consequence of magnetic braking that removes angular momentum from the star on the main sequence. Thus, the age-activity relationship is inherently link to the rotational evolution. As discussed in Section \ref{Chp2_activity-rotation_lit_review}, this has led to studies on the activity-rotation relationship (see e.g. \citealt{Pizzolato_etal_2003,Wright_etal_2011}). The aim of this study was to collect literature values for the rotation period of the sample of stars considered in the two previous studies and investigate the activity-rotation relationship.

\subsection{Determining rotation periods}

There are several possible methods to determine the rotation period of a star; since the rotation periods considered in this work were collected from the literature and use various methods, they are summarised here for convenience.

The first method may be the most commonly used to determine the rotation period for a star - photometric variation in light curves due to starspots. As discussed in Section \ref{Chp1_starspots}, starspots are observed in light curves as periodic variations in the brightness of star as the starspot rotates in and out of view of the observer. To obtain the stellar rotation period, techniques are used to detect periodicities in the light curve. One such technique is the Lomb-Scargle periodogram \citep{Lomb_1976,Scargle_1982}, which uses a Fourier transform to search for periodicities. Typically, a distribution of peaks will appear in the power spectrum and the period with the largest Lomb-Scargle power will correspond to the rotation period. It is a commonly used technique to determine the stellar rotation period (see e.g. \citealt{do_Nascimento_etal_2014,Nielsen_etal_2013}).

Another technique used to determine periodic variations in light curve is the autocorrelation function (ACF); this method determines how correlated the light curve is with itself when offset by a certain time lag. This can be performed for a range of time lags with repeated spot crossings providing peaks in the ACF at time periods associated with the rotation period. An advantage of this method is that the shape of the periodicity in the light curve is not assumed and thus may be more useful in cases where the modulation is not perfectly sinusoidal. This technique is also fairly common in the literature and has been used for large samples of stars from \textit{Kepler} \citep{McQuillan_etal_2014}.

The second method used to determine stellar rotation periods is long term observations of magnetic activity indicators such as the chromospheric emission from the \caII or \Halpha spectral lines and even X-ray luminosity. It is known that these magnetic activity indicators trace active magnetic regions on the surface, thus if an active region crosses the stellar surface multiple times then the rotation period can be determined from the modulation in the magnetic activity indicator. This technique is ideal for stars with more subtle light curve modulation (e.g. slower rotators) as magnetic activity indicators can trace smaller active regions. Example of this method in the literature include \citet{Boro_Saikia_etal_2016,Robertson_etal_2015_GJ176,DeWarf_etal_2010}.

The third method that can be used to determine the rotation period is from asteroseismology. As discussed in Section \ref{Section:intro_ages}, asteroseismology is a valuable tool for determining fundamental parameters through observations of stellar oscillations. However, one aspect not discussed in the previous section is that any departure from solid body rotation will result in a frequency splitting of non-radial oscillation modes; thus the frequency will also depend on the azimuthal order, m. From helioseismology, the analysis of the frequency splitting due to rotation has provided detail about the interior of the Sun and revealed the presence of the tachocline \citep{Spiegel_Zahn_1992} which is instrumental in stellar dynamo theory . However, in order to obtain rotational splitting from asteroseismology, data sets with timescales on the order of years are required. This is due to the amplitude of the frequency splitting which vary from a few micro-Hertz for more massive or younger stars down to a fraction of a micro-Hertz for less massive or older stars. The frequency splitting of p modes in solar-like oscillators are predominantly determined by the rotational profile in the stellar envelope. There is still some debate in the literature about the accuracy of asteroseismic rotation periods (see e.g. \citealt{Barnes_etal_2016_aspect_gyro}), which will be discussed later in this Chapter.

\section{Data and Method}

The sample of stars considered in this analysis comes from the X-ray luminosity and calcium emission studies conducted in Chapters \ref{Chapter3} and \ref{Chapter4}, respectively. Rotation periods for this sample of stars were searched for in the literature using \textit{VizieR} \citep{Ochsenbein_etal_2000}. In addition to this, literature values for the \Rprime activity indicator were collected for the sample of stars from the X-ray luminosity study. The details of the literature values collected and the relevant references for the sample of stars are shown in Appendix \ref{App_I_activity_rotation}.

Since the sample of stars considered in this work is fairly small, comparison samples from previous activity-rotation studies were also used. To compare the stars with values for the \Rprime indicator, the sample of stars from \citet{Metcalfe_etal_2016} were also plotted. Also, to put the asteroseismic samples into context, data was taken from \citet{Baliunas_etal_1996}. For the X-ray sample, data was taken from \citet{Wright_etal_2011} (W11) used to put the sample into context.

As discussed in Section \ref{Chp2_activity-rotation_lit_review}, the activity-rotation relationship is usually calibrated in terms of Rossby number (\Ro). \Ro is defined as the as the rotation period divided by the convective turnover time (\tauc). In this work, \tauc was calculated using the relations from W11 which was derived empirically. By using an empirical method, it assumes that the activity-rotation relationship can be divided into saturated and non-saturated regimes. W11 divided their sample into bins of $B-V$ colour with approximately equal number of stars in each bin. For each mass bin, the activity-rotation relationship was fitted with Equation \ref{Eq:W11_tauc_method_fit_eq} where $(\frac{L_{x}}{L_{BOL}})_{sat}$ and $P_{sat}$ are fitted for each of the mass bins using the $\beta$ value found in W11.

\begin{equation}
    \frac{L_{x}}{L_{BOL}} = 
    \begin{cases}
        C_{B-V}P_{rot}^{\beta} & P_{rot} > P_{sat} \\
        (\frac{L_{x}}{L_{BOL}})_{sat} & P_{rot} < P_{sat}
    \end{cases}
    \label{Eq:W11_tauc_method_fit_eq}
\end{equation}

The fits from each mass bins were used by W11 to determine the colour dependent constant ($C_{B-V}$). The colour dependent constant is defined in Equation \ref{Eq:W11_tauc_method_CBV}, W11 found the colour dependence by setting the scaling constant (C) so that the \tauc value for solar-mass stars matched the values from \citet{Noyes_etal_1984}. This allowed for \tauc to be plotted as a function of colour and/or mass and the best-fitting relationship to be found. It is these best-fitting relationships shown in Equations \ref{Eq:W11_tauc_VK} and \ref{Eq:W11_tauc_mass} that are used to calculate \tauc in this work. In Equation \ref{Eq:W11_tauc_VK}, X is equal to $V-K$ and is valid in the range - $1.1 < V-K < 6.6$.

\begin{equation}
    C_{B-V} = C\tau_{c}^{-\beta}
    \label{Eq:W11_tauc_method_CBV}
\end{equation}

\begin{equation}
    \log \tau_{c} = 
    \begin{cases}
        0.73 + 0.22X & X < 3.5 \\
        -2.16 +1.50X - 0.13X^{2} & X > 3.5
    \end{cases}
    \label{Eq:W11_tauc_VK}
\end{equation}

\begin{equation}
    \log \tau = 1.16 - 1.49\log(M/M_{\odot}) - 0.54\log^{2}(M/M_{\odot})
    \label{Eq:W11_tauc_mass}
\end{equation}

For the majority of samples considered in this analysis, masses were available and \tauc was calculated using Equation \ref{Eq:W11_tauc_mass}. However, for the sample of stars from \citet{Baliunas_etal_1996}, no masses were available. Therefore, Table 3 from \citet{Pecaut_etal_2012} was used to convert $B-V$ values into $V-K$ values and Equation \ref{Eq:W11_tauc_VK} was used to calculate \tauc. Since the table from \citet{Pecaut_etal_2012} is limited to $B-V < 0.77$, the sample from \citet{Baliunas_etal_1996} was also limited to stars that fell into this range.

\section{Results}

\subsection{Chromospheric emission - rotation}
From the calcium emission study, eleven stars had rotation periods found in the literature; a further ten stars from the X-ray emission study had literature values for both the rotation period and the \Rprime activity indicator. 


\subsection{X-ray luminosity - rotation}

\subsection{Rotation - age}

\section{Discussion}


\section{Conclusions}


























